Ecosystem functional response across precipitation extremes in a sagebrush steppe

### 2017-12-22 manuscript/usses_water_ms.Rmd
adding some details on quads and sample sizes

We conducted our precipitation manipulation experiment in a sagebrush steppe community at the USDA, U.S. Sheep Experiment Station (USSES) near Dubois, Idaho (44.2$^{\circ}$ N, 112.1$^{\circ}$ W), 1500 m above sea level.
The plant community is dominated by the shrub \emph{Artemesia tripartita}, \new{the perennial forb \emph{Balsamorhiza sagittata}}, and three perennial bunchgrasses, \emph{Pseudoroegneria spicata}, \emph{Poa secunda}, and \emph{Hesperostipa comata} \new{(see Appendix 1 for rank abundance curves)}.
During the period of our experiment (2011 -- 2015), average mean annual precipitation was 265 mm year$\phantom{}^{-1}$ and mean monthly temperature ranged from -5.2$^{\circ}$C in January to 21.8$^{\circ}$C in July.
Between 1926 and 1932, range scientists at the USSES established 26 permanent 1 m$^2$ quadrats to track vegetation change over time.
In 2007, we relocated 14 of the original quadrats, six of which were inside a large, permanent livestock exclosure.
We use these six plots as control plots (i.e. ambient precipitation) in the experiment described below.
\new{We used the relocated plots as our controls because collecting demographic data is difficult and it is already being collected in these plots for other ongoing studies.}

In spring 2011, we established 16 new 1 m$^2$ plots located in the same exclosure as the six control plots.
We avoided areas on steep hill slopes, areas with greater than 20\% cover of bare rock, and areas with greater than 10\% cover of the shrubs \emph{Purshia tridentata} and/or \emph{Amelanchier utahensis}.
We established the new plots in pairs and randomly assigned each plot in a pair to receive a "drought" or "irrigation" treatment.

>>>>

We conducted our precipitation manipulation experiment in a sagebrush steppe community at the USDA, U.S. Sheep Experiment Station (USSES) near Dubois, Idaho (44.2$^{\circ}$ N, 112.1$^{\circ}$ W), 1500 m above sea level.
The plant community is dominated by the shrub \emph{Artemesia tripartita}, \new{the perennial forb \emph{Balsamorhiza sagittata}}, and three perennial bunchgrasses, \emph{Pseudoroegneria spicata}, \emph{Poa secunda}, and \emph{Hesperostipa comata} \new{(see Appendix 1 for rank abundance curves)}.
During the period of our experiment (2011 -- 2015), average mean annual precipitation was 265 mm year$\phantom{}^{-1}$ and mean monthly temperature ranged from -5.2$^{\circ}$C in January to 21.8$^{\circ}$C in July.
Between 1926 and 1932, range scientists at the USSES established 26 permanent 1 m$^2$ quadrats to track vegetation change over time.
In 2007, we relocated 14 of the original quadrats in permenanent livestock exclosures, \new{which we used} as control plots (i.e. ambient precipitation) in the experiment described below.
\new{We used the relocated plots as our controls because collecting demographic data is difficult and it is already being collected in these plots for other ongoing studies.}

In spring 2011, we established 16 new 1 m$^2$ plots located in the \new{largest exclosure at USSES, which also contained} six of our control plots.
We avoided areas on steep hill slopes, areas with greater than 20\% cover of bare rock, and areas with greater than 10\% cover of the shrubs \emph{Purshia tridentata} and/or \emph{Amelanchier utahensis}.
We established the new plots in pairs and randomly assigned each plot in a pair to receive a "drought" or "irrigation" treatment.
\new{Thus, our experiment consisted of $n=14$ control plots, $n=8$ irrigation plots, and $n=8$ irrigation plots, for a total of 30 plots.}

### 2017-12-22 manuscript/usses_water_ms.tex
adding some details on quads and sample sizes

\definecolor{blue}{rgb}{0,0,0.7} \newcommand{\new}{\textcolor{blue}}

\newcommand\blfootnote[1]{%
  \begingroup
  \renewcommand\thefootnote{}\footnote{#1}%
  \addtocounter{footnote}{-1}%
  \endgroup
}

\blfootnote{*Authors contributed equally to this work.}

@@

\section{INTRODUCTION}\label{introduction}

@@

\section{METHODS}\label{methods}

\subsection{Study Area}\label{study-area}

We conducted our precipitation manipulation experiment in a sagebrush
steppe community at the USDA, U.S. Sheep Experiment Station (USSES) near
Dubois, Idaho (44.2\(^{\circ}\) N, 112.1\(^{\circ}\) W), 1500 m above
sea level. The plant community is dominated by the shrub
\emph{Artemesia tripartita},
\new{the perennial forb \emph{Balsamorhiza sagittata}}, and three
perennial bunchgrasses, \emph{Pseudoroegneria spicata},
\emph{Poa secunda}, and \emph{Hesperostipa comata}
\new{(see Appendix 1 for rank abundance curves)}. During the period of
our experiment (2011 -- 2015), average mean annual precipitation was 265
mm year\(\phantom{}^{-1}\) and mean monthly temperature ranged from
-5.2\(^{\circ}\)C in January to 21.8\(^{\circ}\)C in July. Between 1926
and 1932, range scientists at the USSES established 26 permanent 1
m\(^2\) quadrats to track vegetation change over time. In 2007, we
relocated 14 of the original quadrats, six of which were inside a large,
permanent livestock exclosure. We use these six plots as control plots
(i.e.~ambient precipitation) in the experiment described below.
\new{We used the relocated plots as our controls because collecting demographic data is difficult and it is already being collected in these plots for other ongoing studies.}

In spring 2011, we established 16 new 1 m\(^2\) plots located in the
same exclosure as the six control plots. We avoided areas on steep hill
slopes, areas with greater than 20\% cover of bare rock, and areas with
greater than 10\% cover of the shrubs \emph{Purshia tridentata} and/or
\emph{Amelanchier utahensis}. We established the new plots in pairs and
randomly assigned each plot in a pair to receive a ``drought'' or
``irrigation'' treatment.

\subsection{Precipitation Experiment}\label{precipitation-experiment}

@@

\subsection{Data Collection}\label{data-collection}

@@

\subsection{Data Analysis}\label{data-analysis}

\subsubsection{Ecosystem functional
response}\label{ecosystem-functional-response}

@@

\subsubsection{Sensitivity of ANPP over
time}\label{sensitivity-of-anpp-over-time}

@@

\subsubsection{Community composition over
time}\label{community-composition-over-time}

@@

\subsubsection{Reproducibility}\label{reproducibility}

@@

\section{RESULTS}\label{results}

@@

\section{DISCUSSION}\label{discussion}

@@

\section{CONCLUSIONS}\label{conclusions}

@@

\section{ACKNOWLEDGEMENTS}\label{acknowledgements}

@@

\section{FUNDING}\label{funding}

@@

\section{AUTHOR CONTRIBUTIONS}\label{author-contributions}

@@

\section{SUPPLEMENTAL INFORMATION}\label{supplemental-information}

@@

\section{TABLES}\label{tables}

@@

\section{FIGURES}\label{figures}

@@

\bibliography{/Users/atredenn/Dropbox/Bibliography/usses_water.bib}

>>>>

\definecolor{blue}{rgb}{0,0,0.7}\newcommand{\new}{\textcolor{blue}}

\newcommand\blfootnote[1]{%
  \begingroup
  \renewcommand\thefootnote{}\footnote{#1}%
  \addtocounter{footnote}{-1}%
  \endgroup
}\blfootnote{*Authors contributed equally to this work.}

@@

\hypertarget{introduction}{%
\section{INTRODUCTION}\label{introduction}}

@@

\hypertarget{methods}{%
\section{METHODS}\label{methods}}

\hypertarget{study-area}{%
\subsection{Study Area}\label{study-area}}

We conducted our precipitation manipulation experiment in a sagebrush
steppe community at the USDA, U.S. Sheep Experiment Station (USSES) near
Dubois, Idaho (44.2\(^{\circ}\) N, 112.1\(^{\circ}\) W), 1500 m above
sea level. The plant community is dominated by the shrub
\emph{Artemesia tripartita},
\new{the perennial forb \emph{Balsamorhiza sagittata}}, and three
perennial bunchgrasses, \emph{Pseudoroegneria spicata},
\emph{Poa secunda}, and \emph{Hesperostipa comata}
\new{(see Appendix 1 for rank abundance curves)}. During the period of
our experiment (2011 -- 2015), average mean annual precipitation was 265
mm year\(\phantom{}^{-1}\) and mean monthly temperature ranged from
-5.2\(^{\circ}\)C in January to 21.8\(^{\circ}\)C in July. Between 1926
and 1932, range scientists at the USSES established 26 permanent 1
m\(^2\) quadrats to track vegetation change over time. In 2007, we
relocated 14 of the original quadrats in permenanent livestock
exclosures, \new{which we used} as control plots (i.e.~ambient
precipitation) in the experiment described below.
\new{We used the relocated plots as our controls because collecting demographic data is difficult and it is already being collected in these plots for other ongoing studies.}

In spring 2011, we established 16 new 1 m\(^2\) plots located in the
\new{largest exclosure at USSES, which also contained} six of our
control plots. We avoided areas on steep hill slopes, areas with greater
than 20\% cover of bare rock, and areas with greater than 10\% cover of
the shrubs \emph{Purshia tridentata} and/or
\emph{Amelanchier utahensis}. We established the new plots in pairs and
randomly assigned each plot in a pair to receive a ``drought'' or
``irrigation'' treatment.
\new{Thus, our experiment consisted of $n=14$ control plots, $n=8$ irrigation plots, and $n=8$ irrigation plots, for a total of 30 plots.}

\hypertarget{precipitation-experiment}{%
\subsection{Precipitation Experiment}\label{precipitation-experiment}}

@@

\hypertarget{data-collection}{%
\subsection{Data Collection}\label{data-collection}}

@@

\hypertarget{data-analysis}{%
\subsection{Data Analysis}\label{data-analysis}}

\hypertarget{ecosystem-functional-response}{%
\subsubsection{Ecosystem functional
response}\label{ecosystem-functional-response}}

@@

\hypertarget{sensitivity-of-anpp-over-time}{%
\subsubsection{Sensitivity of ANPP over
time}\label{sensitivity-of-anpp-over-time}}

@@

\hypertarget{community-composition-over-time}{%
\subsubsection{Community composition over
time}\label{community-composition-over-time}}

@@

\hypertarget{reproducibility}{%
\subsubsection{Reproducibility}\label{reproducibility}}

@@

\hypertarget{results}{%
\section{RESULTS}\label{results}}

@@

\hypertarget{discussion}{%
\section{DISCUSSION}\label{discussion}}

@@

\hypertarget{conclusions}{%
\section{CONCLUSIONS}\label{conclusions}}

@@

\hypertarget{acknowledgements}{%
\section{ACKNOWLEDGEMENTS}\label{acknowledgements}}

@@

\hypertarget{funding}{%
\section{FUNDING}\label{funding}}

@@

\hypertarget{author-contributions}{%
\section{AUTHOR CONTRIBUTIONS}\label{author-contributions}}

@@

\hypertarget{supplemental-information}{%
\section{SUPPLEMENTAL INFORMATION}\label{supplemental-information}}

@@

\hypertarget{tables}{%
\section{TABLES}\label{tables}}

@@

\hypertarget{figures}{%
\section{FIGURES}\label{figures}}

@@

\bibliography{/Users/atredenn/Dropbox/Bibliography/usses\_water.bib}
